---
title: "Geospatial Research"
author: "Caitlyn Cai"
date: "2023-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(spData)
library(dplyr)
library(tidyverse)
library(sf)
library(raster)
library(terra)
library(tmap)
library(data.table)
library(sp)
library(maptools)
library(spatstat)
```


# sf for single band
st_read()
read map data, such as .shp
read vector data
# raster for multiple band
read raster data
``raster()``
``brick()``

```{r}
data <- fread("D:/Download/denue_09_2010.csv")
data

data20 <- fread("D:/Download/denue_09_2022_11.csv")
data20
```

```{r}\


```{r}
check01 <- data %>% 
  filter(str_detect(data$nom_estab, regex("ABARROTES|abarroterias", ignore_case = T))) %>% 
  filter(codigo_act != 461110)
check01
reassign01 <- check01 %>% mutate(codigo_act = 461110)

check02 <- data %>% 
  filter(str_detect(data$nom_estab, regex("MINISUPERS|MINI SUPER|MINISUPER", ignore_case = T))) %>% 
  filter(codigo_act != 462112) %>% filter(codigo_act != 461110)
check02
reassign02 <- check02 %>% mutate(codigo_act = 462112)

check03 <- data %>% 
  filter(str_detect(data$nom_estab, regex("polleria|tortilleria", ignore_case = T))) %>% 
  filter(codigo_act != 461190)
check03
reassign03 <- check03 %>% mutate(codigo_act = 461190)

check04 <- data %>% 
  filter(str_detect(data$nom_estab, regex("7 ELEVEN|7ELEVEN", ignore_case = T))) %>% 
  filter(codigo_act != 462112) %>% mutate(codigo_act = 462112) # 0

check05 <- data %>% 
  filter(str_detect(data$nom_estab, ".*\\bOXXO\\b.*")) %>% 
  filter(codigo_act != 462112) %>% mutate(codigo_act = 462112)
check05
reassign05 <- check05 %>% mutate(codigo_act =  462112)

```

```{r}
# Export File.
combine <- rbind(ori_check, reassign01, reassign02, reassign03, reassign05)
write.csv(combine, "D:/Download/clean_week02.csv", row.names=FALSE)

combine
```

# Week 04 03/05/2023  

```{r}
library(sp)
library(maptools)
library(spatstat)
library(sf)
library(foreign)
library(ggplot2)
library(dplyr)
df10 <- read.csv("D:/Download/df10.csv")
df20 <- read.csv("D:/Download/df20.csv")

partial_10 <- dplyr::select(df10, c('longitud','latitud', 'year'))
partial_20 <- dplyr::select(df20, c('longitud','latitud', 'year'))
combine <- rbind(partial_10,filter(partial_20, partial_20$latitud > 19))
# sort(partial_10$longitud, decreasing = FALSE), all correct
sort(partial_20$longitud, decreasing = FALSE)
sort(partial_20$latitud, decreasing = FALSE)

# 13 location points from df20.csv seem to be incorrect.

problem <- filter(df20, df20$latitud < 19)
```


```{r}
write.csv(problem, "D:/Download/13_offpoints.csv", row.names=FALSE)
```


## Write out My shape file  

```{}

WGScoor<- df10
coordinates(WGScoor)=~longitud + latitud
proj4string(WGScoor)<- CRS("+proj=longlat +datum=WGS84")

LLcoor<-spTransform(WGScoor,CRS("+proj=longlat"))
raster::shapefile(LLcoor, "D:/Download/MyShapefile_10.shp")


WGScoor<- df20
coordinates(WGScoor)=~longitud + latitud
proj4string(WGScoor)<- CRS("+proj=longlat +datum=WGS84")

LLcoor<-spTransform(WGScoor,CRS("+proj=longlat"))
raster::shapefile(LLcoor, "D:/Download/MyShapefile_20.shp")
```


## Identify births and deaths

### Attempt 01
```{r}
shp <- st_read("D:/Download/mx_city.shp")
points10 <- st_read("D:/Download/MyShapefile_10.shp")
plot(shp$geometry, main = "Stores in 2010")
plot(points, pch = 1, cex = 0.5,color = "blue", add = T)


shp <- st_read("D:/Download/mx_city.shp")
points20 <- st_read("D:/Download/MyShapefile_20.shp")
plot(shp$geometry, main = "Stores in 2020")
plot(points, pch = 1, cex = 0.5, color = "blue", add = T)
```

## Attempt 02
```{r}
library(mapview)
df10 <- df10 %>% glimpse()
mapview(df10, xcol = "longitud", ycol = "latitud", crs = 4269, grid = FALSE)
mapview(shp)
```


```{r}
mapview(list(shp,points),xcol = "longitud", ycol = "latitud",
        layer.name = c("Mexcian City", "Stores Location"))
```


## Attempt 03
```{r}
df10 <- st_as_sf(df10, coords = c('longitud', 'latitud'))
ggplot(df10) + geom_sf(aes())+ xlim(-99.5,-98.9) + ylim(19.1,19.8)


df20 <- st_as_sf(df20, coords = c('longitud', 'latitud'))
ggplot(df20) + geom_sf(aes()) + xlim(-99.5,-98.9) + ylim(19.1,19.8)
```
