---
title: "Geospatial_MX_Plot"
author: "Caitlyn Cai"
date: "2023-04-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(sf)
library(ggplot2)
library(sp)
library(maptools)
library(spatstat)
library(scales) # load the scales package
library(tmap)
library(RColorBrewer)
```

1) Fill out 2 tables in "Descriptive tab" with clean data.

```{r}
data <- read.csv("D:/Download/df_total_clean.csv")
# birth: only appear in 2020
birth <- data %>% filter(y2010 == 0 & y2020 == 1)

type_birth <- unique(birth$type)
birth_df <- data.frame(type = type_birth, count = NA)
for (i in 1:length(type_birth)) {
  birth_df$count[i] <- count(filter(birth, birth$type == type_birth[i]))$n
}
birth_df

# death: only appear in 2010
death <- data %>% filter(y2010 == 1 & y2020 == 0)

type_death <- unique(death$type)
death_df <- data.frame(type = type_death, count = NA)
for (i in 1:length(type_death)) {
  death_df$count[i] <- count(filter(death, death$type == type_death[i]))$n
}
death_df

# alive: appear in both year

alive <- data %>% filter(y2010 == 1 & y2020 == 1)

type_alive <- unique(alive$type)
alive_df <- data.frame(type = type_alive, count = NA)
for (i in 1:length(type_alive)) {
  alive_df$count[i] <- count(filter(alive, alive$type == type_alive[i]))$n
}
alive_df


df10 <- data %>% filter(y2010 == 1)
type_10 <- unique(df10$type)
ten_df <- data.frame(type = type_10, count = NA, percentage = NA)
for (i in 1:length(type_10)) {
  ten_df$count[i] <- count(filter(df10, df10$type == type_10[i]))$n
  ten_df$percentage[i] <- ten_df$count[i] / count(df10)$n
 }
ten_df

df20 <- data %>% filter(y2020 == 1)
type_20 <- unique(df20$type)
twe_df <- data.frame(type = type_20, count = NA, percentage = NA)
for (i in 1:length(type_20)) {
  twe_df$count[i] <- count(filter(df20, df20$type == type_20[i]))$n
  twe_df$percentage[i] <- twe_df$count[i] / count(df20)$n
 }
twe_df
```



3) Do a choropleth map for each type of establishment.

4) For each establishment category, do location quotients. Estimate location quotient by municipalities for each type of establishment (i.e., number of supermarkets within municipality/total number of supermarkets in the city). Make a map for each type of establishment.  

5) Replicate lab 2 only for "abarrotes" category - do all functions.

6) Only using K function, analyze how much each type of establishment co-locates with itself.

7) Separate categories (abarrotes + specialty) as "traditional" establishments and (convenience + supermarkets) as "modern" establishments. Use cross-K function to analyze how traditional and modern establishments co-locate.


8) Using ALL of the dataset, use cross-K function to analyze how for each type of establishment: 
    a) births co-locate,
    b) deaths co-locate, and finally, 
    c) to analyze how much births from modern establishments co-locate with traditional establishments deaths.

```{r}
data <- read.csv("D:/Download/df_total_clean.csv")
df20 <- data %>% filter(y2020 == 1)
```

```{}

WGScoor<- df20
coordinates(WGScoor)=~longitud + latitud
proj4string(WGScoor)<- CRS("+proj=longlat +datum=WGS84")

LLcoor<-spTransform(WGScoor,CRS("+proj=longlat"))
raster::shapefile(LLcoor, "D:/Download/MyShapefile_20.shp", overwrite = TRUE)
```

```{r}
mx_city	<- readShapeSpatial("D:/Download/mx_city.shp")
mx_pts <- readShapeSpatial("D:/Download/MyShapefile_20.shp")

plot(mx_city,	main= "Mexico City Grocery Stores in Year 2020")
plot(mx_pts, pch=1,cex=0.5,col="blue",add=T)


mx_pt1	<- as.ppp(mx_pts)
# summary(mx_pt1)
# plot(mx_pt1) # need to unmark
mx_pt1only <- unmark(mx_pt1)
plot(mx_pt1only)

mx_bnd <- as.owin(mx_city)
mx2	<- mx_pt1only[mx_bnd]
plot(mx2,main= "Mexicao City Grocery	Stores",cex=0.5,cols="blue")
```
```{r}
# 
qc <- quadratcount(mx2)
plot(qc)
plot(mx_bnd,main= "Mexico City Grocery	Stores in Year 2019")
plot(intensity(qc,image=TRUE),box=FALSE,ribbon=TRUE,add=T)
#plot(mx2,cex=0.5,cols="green",add=T)
```

## Kernel density plot  
```{r}
dd1	<- density(mx2)
summary(dd1)
plot(mx_bnd,main="Mexico City Grocery	Stores in Year 2019")
plot(dd1,box=FALSE,ribbon=TRUE, add = T)
#plot(mx2,cex=0.5,cols="green",add=T)
```
# Minimum squared error

```{r}
bw1	<- bw.diggle(mx2)
plot(bw1)
plot(mx_bnd,main="Mexico City Grocery	Stores in Year 2019	- Min	MSE")
plot(density(mx2,sigma=bw1),box=FALSE,ribbon=TRUE,add=T)
#plot(mx2,cex=0.5,cols="green",add=T)
```

# Maximizing the likelihood  

```{r}
bw2	<- bw.ppl(mx2)
plot(bw2)
plot(mx_bnd,main= "Mexico City Grocery	Stores in Year 2019	- Max	Likelihood")
plot(density(mx2,sigma=bw2),box=FALSE,ribbon=TRUE,add=T)
plot(mx2,cex=0.5,cols="green",add=T)
```
# Cumulative distance function  
```{r}
bw3	<- bw.frac(mx2)
bw3
# plot(bw3)
bw31	<- bw.frac(mx2,f=0.1)
plot(bw31)
plot(mx_bnd,main = "Mexico City Grocery	Stores in Year 2019	- Cum	Distance")
plot(density(mx2,sigma=bw3),box=FALSE,ribbon=TRUE,add=T)
plot(mx2,cex=0.5,cols="green",add=T)
```

# adaptive density  
```{r}
dd2	<- adaptive.density(mx2)

plot(mx_bnd,main="Mexico City Grocery	Stores in Year 2019 - adaptive density")
plot(dd2,box=FALSE,ribbon=TRUE,add=T)
plot(mx2,cex=0.5,cols="green",add=T)

dd3	<- adaptive.density(mx2,f=0.1,nrep=10)

plot(mx_bnd,main="Mexico City Grocery	Stores in Year 2019 - adaptive density")
plot(dd3,box=FALSE,ribbon=TRUE,add=T)
plot(mx2,cex=0.5,cols="green",add=T)
```


# Do a choropleth map for each type of establishment.
### Super

```{r}

stores <-  filter(df20 ,df20$type == "super")
store_counts <- aggregate(cbind(count = rep(1, nrow(stores))) ~ municipio, data = stores, sum)
names(store_counts)[names(store_counts) == "municipio"] <- "muni_name"
store_counts

mx_city$muni_name <- tolower(mx_city$muni_name)
# Merge the data frames by the lowercase "name" column

mx_city_s <- merge(mx_city, store_counts, by = "muni_name")
tm_shape(mx_city_s, title = "Choropleth Map of Supermarket") +
tm_fill("count",
        n = 6,
        style = "quantile",
        palette = "Blues") +
tm_borders(alpha = 0.5) +
tmap_style("classic")


```

### Convience  

```{r}
# Convience
stores <-  filter(df20 ,df20$type == "conv")
store_counts <- aggregate(cbind(count = rep(1, nrow(stores))) ~ municipio, data = stores, sum)
names(store_counts)[names(store_counts) == "municipio"] <- "muni_name"
store_counts

mx_city$muni_name <- tolower(mx_city$muni_name)

# Merge the data frames by the lowercase "name" column

mx_city_s <- merge(mx_city, store_counts, by = "muni_name")
mx_city_s$count[length(mx_city_s$count)] = 0
tm_shape(mx_city_s) + tm_layout(main.title = "Choropleth Map of Convenience Store \nin Year 2019", legend.position = c("left", "top")) + 
tm_fill("count",
        n = 6,
        style = "quantile",
        palette = "-Greens") +
tm_borders(alpha = 0.5) +
tmap_style("classic")
```

### abarrotes  

```{r}
unique(df20$type)
stores <-  filter(df20 ,df20$type == "abarrotes")

store_counts <- aggregate(cbind(count = rep(1, nrow(stores))) ~ municipio, data = stores, sum)
names(store_counts)[names(store_counts) == "municipio"] <- "muni_name"

mx_city$muni_name <- tolower(mx_city$muni_name)
# Merge the data frames by the lowercase "name" column

# create a tmap object
my_map <- tm_shape(mx_city_s) + 
  tm_layout(main.title = "Choropleth Map of Abarrotes \n in Year 2019", 
            main.title.position = "center", 
            legend.position = c("right", "bottom"),
            legend.outside = TRUE,
            legend.outside.position = "right",
            frame = FALSE) +
  tm_fill("count",
          n = 6,
          style = "quantile",
          palette = "YlOrRd") +
  tm_borders(alpha = 0.5) + tmap_style("white")

# save the map as a png file
tmap_save(my_map, "my_map.png")

```
### specialty  

```{r}
stores <-  filter(df20 ,df20$type == "specialty")

store_counts <- aggregate(cbind(count = rep(1, nrow(stores))) ~ municipio, data = stores, sum)
names(store_counts)[names(store_counts) == "municipio"] <- "muni_name"
store_counts

mx_city$muni_name <- tolower(mx_city$muni_name)
# Merge the data frames by the lowercase "name" column

mx_city_s <- merge(mx_city, store_counts, by = "muni_name")
tm_shape(mx_city_s, title = "Choropleth Map of Specialty") +
tm_fill("count",
        n = 6,
        style = "quantile",
        palette = "-Greens") +
tm_borders(alpha = 0.5) +
tmap_style("classic")
```


```{r}
tm_shape(mx_city)+
  tm_polygons(c("DEPENDENCY", "AGED"), style = c("equal", "quantile"), palette = list("Blues", "Greens"))+
  tm_layout(legend.position = c("right", "bottom"))
```


#  For each establishment category, do location quotients.  
Do a choropleth map for each type of establishment.
Make a map for each type of establishment.  
# compare to years, show the shading
```{r}
lq_set <- c()
for (muni in unique(df20$municipio)) {
  result <- df20 %>%
  group_by(type) %>%
  summarize(
    total_estab = n(),
    estab_in_muni = sum(municipio == muni)
  ) %>%
  mutate(
    municipio = muni,
    lq = estab_in_muni / total_estab
  )
  
  lq_set <- rbind(lq_set, result)
}

stores <-  filter(lq_set,lq_set$type == "specialty")

names(stores)[names(stores) == "municipio"] <- "muni_name"
stores

mx_city$muni_name <- tolower(mx_city$muni_name)

mx_city_s <- merge(mx_city, stores, by = "muni_name")

# qtm(mx_city_s, fill ='count', palette = "Blues", title = 'Count Map')

## 
tm_shape(mx_city_s)+
  tm_fill("lq",
          n = 6,
          # style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5) + tmap_style("classic")
```
```{r}
stores <-  filter(lq_set,lq_set$type == "super")

names(stores)[names(stores) == "municipio"] <- "muni_name"
stores

mx_city$muni_name <- tolower(mx_city$muni_name)

mx_city_s <- merge(mx_city, stores, by = "muni_name")

# qtm(mx_city_s, fill ='count', palette = "Blues", title = 'Count Map')



my_map <- tm_shape(mx_city_s) + 
  tm_layout(main.title = "Choropleth Map of Abarrotes \n in Year 2019", 
            main.title.position = "center", 
            legend.position = c("right", "bottom"),
            legend.outside = TRUE,
            legend.outside.position = "right",
            frame = FALSE) +
  tm_fill("lq",
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5) + tmap_style("white")
my_map

```
```{r}
stores <-  filter(lq_set,lq_set$type == "abarrotes")

names(stores)[names(stores) == "municipio"] <- "muni_name"
stores

mx_city$muni_name <- tolower(mx_city$muni_name)
mx_city_s <- merge(mx_city, stores, by = "muni_name")

# qtm(mx_city_s, fill ='count', palette = "Blues", title = 'Count Map')


tm_shape(mx_city_s)+
  tm_fill("lq",
          n = 6,
          # style = "quantile",
          palette = "Reds") +
  tm_borders(alpha = 0.5) + tmap_style("classic")
```
```{r}
stores <-  filter(lq_set,lq_set$type == "conv")

names(stores)[names(stores) == "municipio"] <- "muni_name"
stores

mx_city$muni_name <- tolower(mx_city$muni_name)

mx_city_s <- merge(mx_city, stores, by = "muni_name")



tm_shape(mx_city_s)+
  tm_fill("lq",
          n = 6,
          # style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```

### 三个函数 2019

```{r}
# 6) abarrotes
df_abarrotes <- df20 %>% filter(type == "abarrotes")

df_abarrotes
mx_city	<- readShapeSpatial("D:/Download/mx_city.shp")
abarrotes_pts <- readShapeSpatial("D:/Download/MyShapefile_20_abarrotes.shp")
# plot(mx_city,	main= "Mexico City Abarrotes in Year 2020")
abarrotes_pts	<- as.ppp(abarrotes_pts)
abarrotes_pt1only <- unmark(abarrotes_pts)

mx_bnd <- as.owin(mx_city)
abarrotes_2	<- abarrotes_pt1only[mx_bnd]
plot(abarrotes_2,main= "Mexicao City Abarrotes Storesin Year 2019",cex=0.5,cols="blue")

# par(mfrow = c(3, 1))
gstore	<- Gest(abarrotes_2)
plot(gstore,main= "Mexico City Abarrotes in Year 2019	- G Function")

fstore	<- Fest(abarrotes_2)
plot(fstore,main= "Mexico City Abarrotes in Year 2019	- F Function")

jstore	<- Jest(abarrotes_2)
plot(jstore,main= "Mexico City Abarrotes in Year 2019	- J Function")

png("jstore_2019.png", width = 800, height = 600)

jstore	<- Jest(abarrotes_2)
plot(jstore,main= "Mexico City Abarrotes in Year 2019	- J Function")

# Close the png device and save the plot as a png file
dev.off()
```
### 三个函数 2010

```{r}
# 6) abarrotes
df10_abarrotes <- df10 %>% filter(type == "abarrotes")

df10_abarrotes
mx_city	<- readShapeSpatial("D:/Download/mx_city.shp")
abarrotes_pts <- readShapeSpatial("D:/Download/MyShapefile_10_abarrotes.shp")
abarrotes_pts	<- as.ppp(abarrotes_pts)
abarrotes_pt1only <- unmark(abarrotes_pts)

mx_bnd <- as.owin(mx_city)
abarrotes_2	<- abarrotes_pt1only[mx_bnd]
plot(abarrotes_2,main= "Mexicao City Abarrotes Storesin Year 2019",cex=0.5,cols="blue")

gstore	<- Gest(abarrotes_2)
plot(gstore,main= "Mexico City Abarrotes in Year 2010	- G Function")

fstore	<- Fest(abarrotes_2)
plot(fstore,main= "Mexico City Abarrotes in Year 2010	- F Function")

jstore	<- Jest(abarrotes_2)
plot(jstore,main= "Mexico City Abarrotes in Year 2010	- J Function")
```
7) Separate categories (abarrotes + specialty) as "traditional" establishments and (convenience + supermarkets) as "modern" establishments. Use cross-K function to analyze how traditional and modern establishments co-locate.

```{r}
unique(df20$type)

df20$style <- ifelse(df20$type %in% c("abarrotes", "specialty"), "traditional", "modern")

trad_point	<- readShapeSpatial("D:/Download/MyShapefile_20_trad.shp")
trad_pt1	<- as.ppp(trad_point)
trad_only	<- unmark(trad_pt1)
mx_bnd <- as.owin(mx_city)
trad2	<- trad_only[mx_bnd]
plot(trad2,main= "Traditional Stores",cex=0.5,cols="blue")


mod_point	<- readShapeSpatial("D:/Download/MyShapefile_20_mod.shp")
mod_pt1	<- as.ppp(mod_point)
mod_only	<- unmark(mod_pt1)
mx_bnd <- as.owin(mx_city)
mod2	<- mod_only[mx_bnd]
plot(mod2,main= "Modern Stores",cex=0.5,cols="tomato")


allpt	= superimpose("Traditional"=trad2,"Modern"=mod2,W=mx_bnd)

KCall	<- Kcross(allpt, "Traditional", "Modern")
plot(KCall)
```
```

8) Using ALL of the dataset, use cross-K function to analyze how for each type of establishment: 
    a) births co-locate,
    b) deaths co-locate, 
    c) to analyze how much births from modern establishments co-locate with traditional establishments deaths.
    
```{r}

```

